---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `{rpt}` ([R]{.underline} [p]{.underline}ackage [t]{.underline}emplate)

<!-- badges: start -->

```{r, echo = FALSE, message = FALSE, results='asis'}
library(badger)
cat(
  badge_lifecycle(stage = "experimental", alt = "Lifecycle: experimental"),
  badge_github_actions(alt = "R-CMD-check", action = "R-CMD-check.yaml"),
  badge_codecov(alt = "Codecov test coverage", branch = "main"),
  badger::badge_license()
)
```

<!-- badges: end -->

The goal of `{rpt}` is to provide a template for creating R packages following
[UCD-SERG](https://github.com/UCD-SERG) standards. It includes example
functions, documentation patterns, Quarto-based documentation infrastructure,
and GitHub Actions workflows to help you get started quickly.

## Installation

You can install the development version of `{rpt}` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("UCD-SERG/rpt")
```

## Example

The package ships with two exported functions as starting points.

`example_function()` computes the median of a numeric vector, removing `NA`
values automatically:

```{r example}
library(rpt)
example_function(c(1, 2, 3, 4, 5))
example_function(c(1, NA, 3, 4, 5))
```

`calculate_summary()` returns the mean, median, and standard deviation as a
named list:

```{r summary-example}
calculate_summary(c(1, 2, 3, 4, 5))
```

## Template Features

### Multi-Format Quarto Vignettes and Articles

Vignettes and articles in this template are written in Quarto (`.qmd`) and can
be rendered to multiple output formats from a single source file:

- **HTML** — default format for the documentation website
- **RevealJS slides** — presentation version of every vignette/article
- **DOCX** — Word document version for sharing outside the web

### Slide Break Shortcode

A `slidebreak` shortcode is included and works across all slide deck formats:
RevealJS, PowerPoint (`.pptx`), and Beamer (LaTeX). Use it to insert a new
slide in any format:

``` markdown
{{</* slidebreak */>}}
```

### Mathematical Notation in roxygen2

Function documentation can include inline and display equations using standard
`roxygen2` macros that render correctly in both HTML and PDF help pages:

- `\eqn{x_{0.5}}` for inline math
- `\deqn{m = x_{(n+1)/2}}` for display equations

See `?example_function` for a worked example.

### Quarto Extensions

The following Quarto extensions are preinstalled to enhance documentation:

| Extension | Purpose |
|---|---|
| [`d-morrison/div-anchors`](https://github.com/d-morrison/div-anchors) | Permalink anchors for theorem/proof divs |
| [`d-morrison/equation-anchors`](https://github.com/d-morrison/equation-anchors) | Permalink anchors for numbered equations |
| [`sun123zxy/callouty-theorem`](https://github.com/sun123zxy/callouty-theorem) | Render theorems, definitions, and proofs as callout blocks |

### LaTeX Macro Submodule

A `macros` git submodule (located at `vignettes/macros/`) provides shared LaTeX
macro definitions. Include it in any vignette or article with:

``` markdown
{{</* include macros/macros.qmd */>}}
```

Quarto/pandoc expands the macros at compile time, so MathJax on the website
receives only standard LaTeX.

### Package Structure Visualization

The [Package Function Structure](https://ucd-serg.github.io/rpt/articles/package-structure.html)
article demonstrates how to use
[`foodwebr`](https://lewinfox.com/foodwebr/) to generate dependency graphs for
your package functions.

## Development

### Building the Documentation Site

This package uses [altdoc](https://altdoc.etiennebacher.com/) with [Quarto](https://quarto.org/) to build its documentation site. To build and preview the documentation locally:

```r
# Load the package
pkgload::load_all()

# Render the documentation
altdoc::render_docs()

# Preview the site
altdoc::preview_docs()
```

The documentation is automatically built and deployed to GitHub Pages via GitHub Actions when changes are pushed to the main branch.

## Other R Package Template Options

If you're looking for alternative R package templates, you may also want to consider:

- [r.pkg.template](https://github.com/insightsengineering/r.pkg.template/) - A comprehensive R package template from Insights Engineering

## Code of Conduct

Please note that the `{rpt}` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
