---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `{rpt}` ([R]{.underline} [p]{.underline}ackage [t]{.underline}emplate)

<!-- badges: start -->

```{r, echo = FALSE, message = FALSE, results='asis'}
library(badger)
cat(
  badge_github_actions(alt = "R-CMD-check", action = "R-CMD-check.yaml"),
  badge_codecov(alt = "Codecov test coverage", branch = "main"),
  badge_codefactor(alt = "CodeFactor"),
  badge_cran_release(alt = "CRAN status"),
  badge_lifecycle(stage = "experimental", alt = "Lifecycle: experimental"),
  badge_cran_release(),
  badge_cran_download(type = "grand-total"),
  badge_cran_download(type = "last-month"),
  badge_cran_download(type = "last-week")
)
```

<!-- badges: end -->

The goal of `{rpt}` is to ...

## Installation

You can install the development version of `{rpt}` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("UCD-SERG/rpt")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(rpt)
## basic example code
```

## Development

### Building the pkgdown Site

When building the pkgdown documentation site with `pkgdown::build_site()`, the function needs to download external JavaScript and CSS dependencies from CDNs (Content Delivery Networks). If you're working in a restricted network environment, you'll need to ensure access to the following CDNs:

- `https://cdnjs.cloudflare.com/` - For headroom.js (sticky header functionality)
- `https://cdn.jsdelivr.net/` - For bootstrap-toc.js (table of contents) and other Bootstrap dependencies

Without access to these CDNs, `pkgdown::build_site()` will fail with errors like:

```
Error: cannot open URL 'https://cdnjs.cloudflare.com/ajax/libs/headroom/0.11.0/headroom.min.js'
```

These dependencies are automatically cached after the first download, so subsequent builds will be faster and won't require network access unless the cache is cleared.

## Other R Package Template Options

If you're looking for alternative R package templates, you may also want to consider:

- [r.pkg.template](https://github.com/insightsengineering/r.pkg.template/) - A comprehensive R package template from Insights Engineering

## Code of Conduct

Please note that the `{rpt}` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
