---
title: "Package Function Structure"
author: "UC Davis Seroepidemiology Research Group (UCD-SERG)"
date: "`r Sys.Date()`"
format:
  revealjs:
    output-file: package-structure-revealjs.html
  html: default
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
# Check if we're rendering to a format that supports HTML widgets
is_html_output <- knitr::is_html_output()
```

```{r setup}
#| message: false
library(rpt)
```

## Overview

This article demonstrates how to visualize the structure and dependencies of package functions using the [`foodwebr`](https://lewinfox.com/foodwebr/) package. Understanding function dependencies helps developers:

- Navigate and understand codebases more quickly
- Identify potential refactoring opportunities
- Document architecture and design decisions
- Onboard new contributors more effectively

## What is foodwebr?

The `foodwebr` package creates dependency graphs showing which functions call which other functions. These visualizations are particularly useful for:

- Exploring unfamiliar codebases
- Understanding function relationships
- Identifying central or isolated functions
- Planning refactoring efforts

## Installing foodwebr

If you don't have `foodwebr` installed, you can install it from CRAN:

```{r}
#| eval: false
install.packages("foodwebr")
```

Or from GitHub for the latest development version:

```{r}
#| eval: false
devtools::install_github("lewinfox/foodwebr")
```

## Visualizing Package Structure

### Basic Usage

The `foodweb()` function analyzes function dependencies. When examining a package, you typically want to look at specific functions or the entire package namespace:

```{r}
#| message: false
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE)
library(foodwebr)

# Create a foodweb for the example_function
fw <- foodweb(FUN = example_function)
fw
```

### Plotting the Dependency Graph

The `plot()` method creates an interactive visualization using DiagrammeR:

```{r}
#| label: fig-basic-foodweb
#| fig-cap: "Basic function dependency graph for example_function"
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE) && is_html_output
plot(fw)
```

::: {.callout-note}
## Understanding the Graph

- **Nodes** represent individual functions
- **Edges** (arrows) show function calls
- An arrow from A to B means "A calls B"
- Isolated nodes indicate functions with no dependencies or dependents
:::

### Filtering Options

By default, `foodwebr` filters to show only functions directly related to the specified function. You can control this behavior:

```{r}
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE)
# Show only connected functions (default)
fw_filtered <- foodweb(FUN = example_function, filter = TRUE)

# Show all functions in the environment
fw_all <- foodweb(FUN = example_function, filter = FALSE)
```

### Examining the Entire Package

To see all exported and internal functions in the package:

```{r}
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE)
# Get the package namespace environment
pkg_env <- asNamespace("rpt")

# Create foodweb for entire package
fw_pkg <- foodweb(env = pkg_env)
fw_pkg
```

```{r}
#| label: fig-package-foodweb
#| fig-cap: "Complete dependency graph for all package functions"
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE) && is_html_output
plot(fw_pkg)
```

## Analyzing Dependencies

### Understanding Function Roles

From the dependency graph, you can identify different types of functions:

- **Leaf functions**: Functions that don't call any other package functions (only base R or external packages)
- **Root functions**: Functions that are called by many others but don't call package functions themselves
- **Hub functions**: Functions that both call and are called by many functions
- **Isolated functions**: Functions with no connections to others

### Example Analysis

For our package:

```{r}
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE)
# Convert to text representation for inspection
cat(as.character(fw_pkg))
```

::: {.callout-tip}
## Dependency Patterns

In well-designed packages, you typically see:

1. **Clear hierarchy**: Lower-level utility functions are called by higher-level API functions
2. **Limited cross-dependencies**: Functions in the same "layer" don't call each other excessively
3. **Focused functions**: Each function has a clear purpose with limited dependencies

In the `rpt` package, we can see this pattern:

- `validate_input()` is a low-level helper (no package dependencies)
- `clean_data()` calls `validate_input()` 
- `calculate_statistic()` calls `clean_data()`
- `format_result()` is a low-level formatter (no package dependencies)
- `example_function()` is a high-level API that calls `calculate_statistic()` and `format_result()`
- `calculate_summary()` is another high-level API that calls `clean_data()` and `example_function()`

This creates a clear layered architecture with well-separated concerns.
:::

## Package Architecture

### Current Structure

Let's examine the actual structure of the `rpt` package:

```{r}
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE)
# List all functions in the package
pkg_functions <- ls(asNamespace("rpt"))
cat("Package functions:\n")
cat(paste("-", sort(pkg_functions), collapse = "\n"))
```

The dependency relationships are:

- **Exported functions** (user-facing API):
  - `example_function()`: Calls `calculate_statistic()` → `format_result()`
  - `calculate_summary()`: Calls `clean_data()` → `example_function()`

- **Internal functions** (implementation details):
  - `validate_input()`: No package dependencies (leaf function)
  - `clean_data()`: Calls `validate_input()`
  - `calculate_statistic()`: Calls `clean_data()`
  - `format_result()`: No package dependencies (leaf function)

This architecture demonstrates good separation of concerns with clear data flow from validation through cleaning to calculation and formatting.

## Advanced Usage

### Using with tidygraph

The `tidygraph` package provides powerful tools for graph analysis. You can convert a `foodweb` object to work with these tools:

```{r}
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE) && requireNamespace("tidygraph", quietly = TRUE)
if (requireNamespace("tidygraph", quietly = TRUE)) {
  tg <- tidygraph::as_tbl_graph(fw_pkg)
  print(tg)
}
```

### Customizing Visualizations

You can pass additional arguments to customize the graph appearance:

```{r}
#| eval: false
# These arguments are passed to DiagrammeR::grViz()
plot(fw_pkg, 
     width = 800, 
     height = 600)
```

### Exporting as Text

Get the graphviz DOT representation:

```{r}
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE)
# As text
foodweb_text <- foodweb(env = asNamespace("rpt"), as.text = TRUE)
cat(foodweb_text)
```

## Practical Applications

### Code Review

Use dependency graphs during code review to:

- Verify that new functions follow existing architectural patterns
- Identify unexpected dependencies
- Ensure proper separation of concerns
- Detect circular dependencies

### Refactoring

Before refactoring:

1. Generate a dependency graph of the affected area
2. Identify all functions that will be impacted
3. Plan changes to minimize ripple effects
4. After refactoring, regenerate the graph to verify improvements

### Documentation

Include dependency graphs in:

- Developer documentation
- Architecture decision records
- Package vignettes (like this one!)
- README files for complex packages

::: {.callout-important}
## Keeping Documentation Updated

When adding new functions or modifying dependencies, regenerate the foodweb graphs to keep documentation current. Consider adding automated checks in your CI/CD pipeline.
:::

## Exploring Specific Functions

You can focus on specific functions to understand their dependencies:

### Example Function Dependencies

```{r}
#| label: fig-example-function
#| fig-cap: "Dependency graph for example_function showing all related functions"
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE) && is_html_output
fw_example <- foodweb(FUN = example_function)
plot(fw_example)
```

### Summary Function Dependencies

```{r}
#| label: fig-summary-function
#| fig-cap: "Dependency graph for calculate_summary showing its relationships"
#| eval: !expr requireNamespace("foodwebr", quietly = TRUE) && is_html_output
fw_summary <- foodweb(FUN = calculate_summary)
plot(fw_summary)
```

These focused views help understand:

- What functions a specific function depends on (its dependencies)
- What functions depend on a specific function (its dependents)
- The complete call chain for a given function

## Best Practices

When designing package architecture:

1. **Minimize dependencies**: Each function should have a clear, focused purpose
2. **Avoid circular dependencies**: A should not call B if B calls A
3. **Layer your functions**: Separate low-level utilities from high-level APIs
4. **Document dependencies**: Use `foodwebr` graphs in your documentation
5. **Review regularly**: Check dependency graphs during code review

## Learn More

- **foodwebr documentation**: <https://lewinfox.com/foodwebr/>
- **Package website**: <https://ucd-serg.github.io/rpt/>
- **GitHub repository**: <https://github.com/UCD-SERG/rpt>
- **DiagrammeR**: <https://rich-iannone.github.io/DiagrammeR/>
- **tidygraph**: <https://tidygraph.data-imaginist.com/>

## Summary

This article demonstrated how to use `foodwebr` to visualize and understand package function dependencies. Key takeaways:

- `foodwebr` creates interactive dependency graphs
- Filtering options help focus on specific functions or show entire packages
- Dependency graphs aid in code review, refactoring, and documentation
- Integration with `tidygraph` enables advanced graph analysis
- Regular visualization helps maintain clean architecture

The `rpt` package demonstrates a clean layered architecture with:
- Low-level utility functions (`validate_input()`, `format_result()`)
- Mid-level processing functions (`clean_data()`, `calculate_statistic()`)
- High-level API functions (`example_function()`, `calculate_summary()`)

This structure makes the code easier to understand, test, and maintain.

## References

::: {#refs}
:::
